---
title: "R Programming Vectors"
author: "Krisztina Gyalai"
date: "24/01/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Vectors:**

* Basic data object in R. 
* Can contain the same or different data types. 
* Vector with same data type: **atomic vector**
* Types: character, integer, double, complex, logical, raw

### 1.Creating empty vectors

```{r}
# creating empty vector with c() 
# output = 'NULL'
vector_1 <- c()
#creating empty vector with vector()
#output ='logical (empty)'
vector_2 <- vector()
#creating vector by giving the name to variable and passing NULL into it
#output 'NULL'
vector_3 <- NULL
#using numeric function 
#creates numeric vector output = 'numeric (empty)'
vector_4 <- numeric()
#using the rep() function
#output = 'NULL'
vector_5 <- rep()
#creating an empty nested vector
#output = 'NULL' 
vector_6 <- c(c(),c())
```

### 2.Adding elements to the vector 

```{r}
#using the range (:) operator 
# need to add a start and end value 
#add numbers from 10 to 20 to vector_1
vector_1 <- 10:20
print(vector_1)
```

```{r}
#assigning values to the vector which was empty
print(vector_2)
vector_2 <- c(TRUE,FALSE,FALSE,FALSE,FALSE)
print(vector_2)
```

```{r}
#using index [] operator 
# index value is a location where the particular element is located 
# data value which added to this particular location 
#insert number 100 in location 1 to vector_1 
vector_1[1] <- 100
print(vector_1)
```
```{r}
#insert 0.1 to location 5
vector_1[5] <- 0.1
print(vector_1)

```

Adding float to the integer vector changed all data types inside the vector to float. 

```{r}
#using append() function
#add 200 to vector_1
vector_1 <- append(vector_1,200)
print(vector_1)
```
```{r}
#add multiple values with append c()
#add 300,400,500 to vector_1
vector_1 <- append(vector_1, c(300,400,500))
print(vector_1)
```

### 3.Find Sum, Mean and Product of a vector

In case if there a single vector is specified then the operation will be performed over individual elements. 


```{r}
#create a vector, find the mean include NA in the calculation 
a <- c(5,5,5,5,5)
print(mean(a, na.rm = FALSE))
#create a vector include NA value
b <- c(5,5,5,NA,5)
#save it as a numeric vector
as.numeric(b)
#ignore NA on mean calculation 
print(mean(b,na.rm=TRUE))
#don't ignore NA
c <- b
print(mean(c))
#sum
print(sum(a, na.rm = FALSE))
#product
prod(a)
#with NA s 
prod(b)
#ignore NA's
prod(b, na.rm = TRUE)
```
### 4.Adding, dividing and multiplying vectors by other vector 

```{r}
#add
vector_a <- c(10,20,30)
vector_b <- c(20,20,20)
cat("a:", vector_a,"\nb:",vector_b)
vector_c <- vector_a + vector_b
cat("\na+b:", vector_c)
#divide
vector_d <- vector_a/vector_b
cat("\na/b:",vector_d)
#multiple
vector_e <- vector_a * vector_b 
cat("\na*b:", vector_e)
```
### 5.Find the minimum and maximum element of a vector 

```{r}
#create a vector
a <- as.numeric(c(7.23, 5.88, 9.23, 12.99, 23.1))
cat("vector:", a)
#find minimum
cat("\nMinimum:",min(a))
#find maximum
cat("\nMaximum:",max(a))
```
### 6.Sorting the elements of the vector

```{r}
x <- c(1,5,89,34,55,62,2,45)
print(x)
#sort increasing
print(sort(x))
#descending 
print(sort(x, decreasing = TRUE))

```
### 7 Check wheather the vector contain a given element 

```{r}
print(x)
#check that vector x has the element 5 in it
print(is.element(5,x))
#check that vector x has the element 6 
print(is.element(6, x))
```
### 8. Finding the last element in the vector 

```{r}
print(x)
#find the last element in the vector
print(tail(x, n = 1))
#find the last three elements in the vector
print(tail(x, n=3 ))
```
### 9. Finding the 'n' highest and 'n' lowest number in the vector

```{r}
print(x)
#find the length of the vector
l = length(x)
cat("\nThe Length of the vector is =",l,"\n")
#if we sort the vector in increasing order we know the last number will be the highest
# sort vector
sort(x)
#vector is 8 element long
#the 8th number is the highest 
#to find the second highest number
sorted_x <- sort(x)
#second highest number
cat("\nThe second highest number is:", sorted_x[l-1])
#find the second lowest if we know the lowest number is in location 1
cat("\nThe second lowest number is:",sorted_x[l-6])
```
### 10.Extract every 'nth' element of the vector

```{r}
#create a vector 
y <- c(50:100)
print(y)
#extract every 3th element 
#sequence starting with first element until the last element of the vector jump with 3
n <- y[seq(1, length(y),3)]
print(n)

```
### 11.Find the most common element of multiple vectors

```{r}
fruits_1 <- c("apple","banana","pear")
fruits_2 <- c("pineapple","apple","wine apple")
fruits_3 <- c("apple", "kiwi", "banana")
cat("1:",fruits_1,"\n2:",fruits_2,"\n3:",fruits_3,"\n")
#find the common element
cat("most common element:",intersect(intersect(fruits_1,fruits_2),fruits_3))
```
```{r}
#there are two most common elements : both appear as a result - tie between 
fruits_1 <- c("apple","banana","pear")
fruits_2 <- c("pineapple","apple","banana")
fruits_3 <- c("apple", "kiwi", "banana")
cat("1:",fruits_1,"\n2:",fruits_2,"\n3:",fruits_3,"\n")
#find the common element
cat("most common element:",intersect(intersect(fruits_1,fruits_2),fruits_3))
```
### 12. Find unique values in a vector (distinct elements) 

```{r}
fruits <- c("pineapple", "apple","apple", "banana","pineapple")
print(fruits)
unique(fruits)
```

### 13. Find elements in the vaector that distinct from the other vector 

```{r}
#given two vector
fruits_1 <- c("apple","banana","pear")
fruits_2 <- c("pineapple","apple","banana")
cat("\n1:",fruits_1,"\n2:",fruits_2)
#find the difference 
cat("\ndifference:", setdiff(fruits_1,fruits_2))

```

### 14.Reverse the order of a vector

```{r}
print(fruits)
cat("\nreversed:",rev(fruits))
```
### 15.Concatanate elements of the vector

```{r}
print(fruits)
#merge words together
concat <- paste(fruits, collapse = "")
print(concat)
```

### 16.Count the number of elements of the vector within the given range

```{r}
#create a vector for numbers 1 to 10
x <- c(1:10)
print(x)
#check how many numbers are between 3 and 7
in_between <- sum( x > 3 & x < 7)
print(in_between)
```
### 17.combine two vectors 

```{r}
x <- c("apple","carrot","potato","sugar")
y <- c(1,9,4,7)
cat("\nx:",x, "\ny:",y)
#combine the two vectors vertically
new <- cbind(x,y)
print(new)
#combine vertically
new_2 <- rbind(x,y)
cat("\n",new_2)
```

### 18. Creating vector using the seq() function 

```{r}
#create a vector using seq(), start, end, step 
sequence <- seq(3, 12, by = 0.5)
print(sequence)
#specify the length - how many segment you want to cut the numbers in between 
sequence_2 <- seq(3, 12, length.out = 13)
print(sequence_2)

```

